
```{r}
#| label: fit-models
#| include: false

# EKG Feature ~ Trimester
qrs_preg_lmer <- lmerTest::lmer(
  QRS_Duration ~ Trimester + (1|PregKey),
  data = dat
)

qtc_preg_lmer <- lmerTest::lmer(
  QTc_Bazett ~ Trimester + (1|PregKey),
  data = dat
)

pr_preg_lmer <- lmerTest::lmer(
  PR_Interval ~ Trimester + (1|PregKey),
  data = dat
)

vent_preg_lmer <- lmerTest::lmer(
  VentRate ~ Trimester + (1|PregKey),
  data = dat
)

trimester_ekg_results <- dplyr::left_join(
  process_ekg_lmer(qrs_preg_lmer) |> 
    dplyr::rename(`QRS Duration` = Estimate),
  process_ekg_lmer(qtc_preg_lmer) |> 
    dplyr::rename(`QTc Interval` = Estimate),
  by = ("Predictor" = "Predictor")
) |>
  dplyr::left_join(
    process_ekg_lmer(pr_preg_lmer) |> 
      dplyr::rename(`PR Interval` = Estimate),
    by = ("Predictor" = "Predictor")
  ) |>
  dplyr::left_join(
    process_ekg_lmer(vent_preg_lmer) |> 
      dplyr::rename(`Heart Rate` = Estimate),
    by = ("Predictor" = "Predictor")
  )

ekg_tri_model <- list(
  "PR Interval" = pr_preg_lmer, 
  "QRS Duration" = qrs_preg_lmer, 
  "QTc Interval" = qtc_preg_lmer, 
  "Heart Rate" = vent_preg_lmer
)

trimester_ekg_marginals <- dplyr::bind_cols(
  lapply(
    1:4,
    function(x) {
      ggeffects::predict_response(
        ekg_tri_model[[x]], terms = "Trimester", margin = "mean_reference"
      ) |> 
        tibble::as_tibble() |> 
        # round to 2 decimals
        dplyr::mutate(
          dplyr::across(
            .cols = where(is.numeric),
            ~ .x |> round(2)
          )
        ) |> 
        # combine ci into single column
        tidyr::unite("ci", conf.low, conf.high, sep = ", ") |> 
        # combine est and ci into single column
        tidyr::unite(
          col = "tmp", 
          predicted, ci, 
          sep = " ("
        ) |> 
        dplyr::mutate(
          tmp = paste0(tmp, ")")
        ) |> 
        # rename columns
        dplyr::rename(
          "Trimester" = x,
          !!names(ekg_tri_model)[x] := tmp
        ) |> 
        # remove unnecessary columns
        dplyr::select(-c(std.error, group))
    }
  )
) |> 
  dplyr::rename(
    Trimester = `Trimester...1`
  ) |> 
  dplyr::select(
    -c(`Trimester...3`, `Trimester...5`, `Trimester...7`)
  )
  
  

# multivariable model: preg = ekg pred
preg_ekg_lmer <- lme4::glmer(
  Pregnant ~ PR_Interval + QRS_Duration + QTc_Bazett + VentRate + 
    (1|PregKey),
  data = dat,
  family = binomial
)

# process results
tmp <- summary(preg_ekg_lmer)$coefficients |>
  tibble::as_tibble(rownames = NA) |>
  tibble::rownames_to_column(var = "Variable")
est <- gtsummary::style_sigfig(
  x = exp(tmp$Estimate), 
  digits = 3
)
lower <- gtsummary::style_sigfig(
  x = exp(tmp$Estimate - (1.96 * tmp$`Std. Error`)), 
  digits = 3
)
upper <- gtsummary::style_sigfig(
  x = exp(tmp$Estimate + (1.96 * tmp$`Std. Error`)), 
  digits = 3
)
p_vals <- gtsummary::style_pvalue(tmp$`Pr(>|z|)`)

# combine
preg_ekg_results <- data.frame(
  Predictors = c("Intercept", "PR Interval", "QRS Duration", 
                 "QTc Interval", "Heart Rate"),
  `OR (95% CI)` = paste0(est, " (", lower, ", ", upper, ")"),
  `p-value` = p_vals,
  Explanation = c(
    " ",
    # PR Interval
    paste0("A 10 unit increase in the PR Interval is associated with a ",
           "2.6% decrease in the odds of being pregnant",
           ", controlling for the other ECG characteristics."),
    # QRS Duration
    paste0("A 10 unit increase in the QRS duration is associated with a ",
           "4.4% decrease in the odds of being pregnant",
           ", controlling for the other ECG characteristics."),
    # QTc Interval
    paste0("A 10 unit increase in the QTc Interval is associated with a ",
           "0.1% decrease in the odds of being pregnant",
           ", controlling for the other ECG characteristics."),
    # Heart Rate
    paste0("A 10bpm increase in the heart rate is associated with a ",
           "7% increase in the odds of being pregnant",
           ", controlling for the other ECG characteristics.")
  ),
  check.names = FALSE
) |>
  # remove Intercept
  dplyr::filter(
    Predictors != "Intercept"
  )

# Trend test for EKG features across trimesters
ekg_trend_results <- process_ekg_trends(
  data = dat,
  ekg_vars = c("PR_Interval", "QRS_Duration", "QTc_Bazett", "VentRate"),
  trimesters = c(0, 1, 2, 3, 4)
)
```


After excluding patients with pre-existing myocardial infarction, 
heart failure, heart valve disorders, atrial fibrillation and other arrhythmias, 
along with those who developed hypertensive disorders of pregnancy, 
`r extract_n_ekg(dat)` ECGs from `r extract_n_subj(dat)` patients were 
included: `r n_ekg_not_preg` of ECGs (`r prop_ekg_not_preg`) were obtained 
outside of pregnancy or the 6 months postpartum period and `r n_ekg_preg` of 
ECGs (`r prop_ekg_preg`) during pregnancy or 2 weeks postpartum, 
`r n_ekg_tri4` ECGs were during the 2 weeks postpartum period. 
In a multiple logistic mixed effect model, a higher heart rate were 
associated the pregnancy state (@tbl-preg-ekg)). 
The shorter PR interval was associated with the 2nd and 3rd trimesters, 
while higher heart rate was associated with the third trimester, 
(@tbl-ekg-trimester and @tbl-ekg-trimester-pred).  

::: {#tbl-preg-ekg}

```{r}
flextable::flextable(preg_ekg_results) |>
  flextable::theme_vanilla() |>
  flextable::fontsize(
    i = c(1:4),
    j = c(1, 2),
    size = 10
  ) |>
  flextable::width(
    j = 1, width = 1.75
  ) |>
  flextable::width(
    j = 2, width = 2.25
  ) |>
  flextable::width(
    j = 4,
    width = 4
  )
```
Summary of ECG Features Associated with Pregnancy Compared to Non-pregnant State.
:::


::: {#tbl-ekg-trimester}

```{r}
flextable::flextable(trimester_ekg_results) |>
  flextable::theme_vanilla() |>
  flextable::fontsize(
    i = c(1:6),
    j = c(1:5),
    size = 10
  ) |>
  flextable::width(
    j = 1, width = 1.2
  ) |>
  flextable::width(
    j = c(2:5), width = 1.6
  ) |>
  # manually select cells to bold
  flextable::bold(
    i = 3,
    j = 4
  ) |>
  flextable::bold(
    i = 4,
    j = c(4, 5)
  ) |>
  flextable::bold(
    i = 5,
    j = c(4, 5)
  ) 
```
Results of linear mixed effect models: ECG Features Associated with Pregnancy 
Trimesters compared to period before pregnancy (Trimester 0)
:::

::: {#tbl-ekg-trimester-pred}

```{r}
trimester_ekg_marginals |> 
  flextable::flextable() |> 
  flextable::theme_vanilla() |> 
  flextable::autofit() |> 
  # manually select cells to bold
  flextable::bold(
    i = 3,
    j = 2
  ) |>
  flextable::bold(
    i = 4,
    j = c(2, 5)
  ) |>
  flextable::bold(
    i = 5,
    j = c(2, 5)
  ) 
```
Marginal estimates of ECG Features by Pregnancy Trimesters.
:::

## Trend Tests

To formally test whether ECG features change systematically across pregnancy 
trimesters, we performed trend tests using both non-parametric 
(Jonckheere-Terpstra) and parametric (polynomial contrasts) methods. The 
Jonckheere-Terpstra test is particularly useful as it does not assume a 
linear trend and can detect monotonic changes over time (@tbl-trend-tests).

::: {#tbl-trend-tests}

```{r}
flextable::flextable(ekg_trend_results) |>
  flextable::theme_vanilla() |>
  flextable::autofit() |>
  flextable::fontsize(
    size = 9
  ) |>
  flextable::set_header_labels(
    EKG_Feature = "ECG Feature",
    N = "N",
    JT_p_value = "JT p-value",
    Linear_p = "Linear p",
    Quadratic_p = "Quadratic p",
    Interpretation = "Interpretation"
  )
```
Trend test results for ECG features from pre-pregnancy (Trimester 0) through 
early postpartum (Trimester 4, ≤2 weeks postpartum). JT = Jonckheere-Terpstra 
non-parametric trend test; Linear and Quadratic p-values from polynomial contrasts.
:::
